## 231211 - EDA
```{r, setup, include = FALSE}
knitr::opts_knit$set(root.dir = '~/rna/autoregulation/g4/')
```
I have G4s from CDS, 5' UTR and 3' UTR regions 
Disease causing regions? overlaps with eCLIP BAMs? 

```{r}
library(tidyr)

cds <- read.table("cds.GRCh38.g4.tsv", header = FALSE, 
                  col.names = c("annot", "id", "seq"))
utr5 <- read.table("utr5.GRCh38.g4.tsv", header = FALSE, 
                   col.names = c("annot", "id", "seq"))
utr3 <- read.table("utr3.GRCh38.g4.tsv", header = FALSE, 
                   col.names = c("annot", "id", "seq"))

# head(cds)
# typeof(cds)
```

### Basic info 
```{r}
cat("G4-containing exons in CDS: ", dim(cds)[1], "\n")
cat("G4-containing exons in 5' UTR:", dim(utr5)[1], "\n")
cat("G4-containing exons in 3' UTR:", dim(utr3)[1], "\n")
```
There are a lot more in 3' UTR - is that significant?  
- Exploring mRNA 3′-UTR G-quadruplexes: evidence of roles in both alternative polyadenylation and mRNA shortening  
  - https://pubmed.ncbi.nlm.nih.gov/23609544  

### Tidy up dataframes so info is more easily accessible 
```{r}
cds <- separate(cds, col = "annot", 
                into = c("chr", "pos", "str"), 
                sep = ":")
cds <- separate(cds, col = "pos", 
                into = c("start", "stop"), 
                sep = "-")

utr5 <- separate(utr5, col = "annot", 
                into = c("chr", "pos", "str"), 
                sep = ":")
utr5 <- separate(utr5, col = "pos", 
                into = c("start", "stop"), 
                sep = "-")

utr3 <- separate(utr3, col = "annot", 
                into = c("chr", "pos", "str"), 
                sep = ":")
utr3 <- separate(utr3, col = "pos", 
                into = c("start", "stop"), 
                sep = "-") 
```

### How many in each chrom? 
```{r}
cds.regions <- table(cds$chr)
sort(cds.regions, decreasing = TRUE)

utr5.regions <- table(utr5$chr)
sort(utr5.regions, decreasing = TRUE)

utr3.regions <- table(utr3$chr)
sort(utr3.regions, decreasing = TRUE)
```

### Find eCLIP peak overlaps with G4 regions
https://research.stowers.org/cws/CompGenomics/Tutorial/GRanges/guide.html 
Idea:
1. get BED files from eCLIP data available in lab proj space
1. get ranges data from my G4 files 
1. convert all to GRanges objects 
1. find intersections? 
```{r}
library(GenomicRanges)

```