## 231211 - EDA

```{r, setup, include = FALSE}
knitr::opts_knit$set(root.dir = '~/rna/autoregulation/g4/')
```

I have G4s from CDS, 5' UTR and 3' UTR regions Disease causing regions? overlaps with eCLIP BAMs?

```{r}
library(tidyr)

cds <- read.table("cds.GRCh38.g4.tsv", header = FALSE, 
                  col.names = c("annot", "id", "seq"))
utr5 <- read.table("utr5.GRCh38.g4.tsv", header = FALSE, 
                   col.names = c("annot", "id", "seq"))
utr3 <- read.table("utr3.GRCh38.g4.tsv", header = FALSE, 
                   col.names = c("annot", "id", "seq"))

# head(cds)
# typeof(cds)
```

### Basic info

```{r}
cat("G4-containing exons in CDS: ", dim(cds)[1], "\n")
cat("G4-containing exons in 5' UTR:", dim(utr5)[1], "\n")
cat("G4-containing exons in 3' UTR:", dim(utr3)[1], "\n")
```

-   There are a lot more in 3' UTR - is that significant?

-   Exploring mRNA 3′-UTR G-quadruplexes: evidence of roles in both alternative polyadenylation and mRNA shortening

    -   https://pubmed.ncbi.nlm.nih.gov/23609544

-   Conversation with Dan: when we look at G4 frequency in each region, 5' UTR should have the highest frequency (not in raw counts - the other regions will have more in raw numbers, simply because they are longer regions)

### Tidy up dataframes so info is more easily accessible

```{r}
cds <- separate(cds, col = "annot", 
                into = c("chr", "pos", "str"), 
                sep = ":")
cds <- separate(cds, col = "pos", 
                into = c("start", "stop"), 
                sep = "-")

utr5 <- separate(utr5, col = "annot", 
                into = c("chr", "pos", "str"), 
                sep = ":")
utr5 <- separate(utr5, col = "pos", 
                into = c("start", "stop"), 
                sep = "-")

utr3 <- separate(utr3, col = "annot", 
                into = c("chr", "pos", "str"), 
                sep = ":")
utr3 <- separate(utr3, col = "pos", 
                into = c("start", "stop"), 
                sep = "-") 
```

### How many in each chrom?

```{r}
cds.regions <- table(cds$chr)
sort(cds.regions, decreasing = TRUE)

utr5.regions <- table(utr5$chr)
sort(utr5.regions, decreasing = TRUE)

utr3.regions <- table(utr3$chr)
sort(utr3.regions, decreasing = TRUE)
```

### Find eCLIP peak overlaps with G4 regions

<https://research.stowers.org/cws/CompGenomics/Tutorial/GRanges/guide.html>

Idea:

1.  get BED files from eCLIP data available in lab proj space

2.  get ranges data from my G4 files

3.  convert all to GRanges objects

4.  find intersections?

```{r}
library(GenomicRanges)

```
